---
title: "Exporting Radial Metrics"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Exporting Radial Metrics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  error = TRUE
)
```

```{r setup}
library(SeaSondeR)
```

# Exporting Radial Metrics in SeaSondeR

This vignette demonstrates how to export radial metrics from HF-Radar spectra using the SeaSondeR package. The exported radial metrics table contains key parameters (e.g. DOAs, FOR limits, signal strengths) that can be used with tools such as Sea Display.

## Example

```{r}
# Set sample file paths
cs_file <- system.file("css_data/CSS_TORA_24_04_04_0700.cs", package = "SeaSondeR")

apm_file <- system.file("css_data/MeasPattern.txt", package = "SeaSondeR")

# Read the antenna pattern file to create a SeaSondeRAPM object
apm_obj <- seasonder_readSeaSondeRAPMFile(apm_file)

# Create a SeaSondeRCS object from a spectral file
cs_obj <- seasonder_createSeaSondeRCS(cs_file, seasonder_apm_object = apm_obj)

# Compute first-order regions and run the MUSIC algorithm if needed:

cs_obj <- seasonder_computeFORs(cs_obj, FOR_control = list(nsm = 2, fdown = 10^(10/10), flim = 10^(20/10),
                                                            noisefact = 10^(6/10), currmax = 2,
                                                            reject_distant_bragg = TRUE, reject_noise_ionospheric = FALSE,
                                                            reject_noise_ionospheric_threshold = 0))
 cs_obj <- seasonder_runMUSICInFOR(cs_obj)

# Define a land mask (AngSeg) to remove onshore data
AngSeg <- purrr::list_c(lapply(45:61, function(i) list(c(i, 313, 360), c(i, 0, 31))))

# Export the radial metrics in LLUV format
radial_metrics <- seasonder_exportLLUVRadialMetrics(cs_obj,
                                                    LLUV_path = "CSS_TORA_24_04_04_0700.ruv",
                                                    AngSeg = AngSeg)

# Display the first few rows of the exported radial metrics
head(radial_metrics)
```

The example above exports the radial metrics into a file (LLUV format) that can later be reviewed using visualization applications such as Sea Display.

## Description of the Radial Metrics Table

Below are the columns of the exported Radial Metrics table in SeaSondeR.

| Column               | Description                                                                                              |
| -------------------- | -------------------------------------------------------------------------------------------------------- |
| LOND, LATD           | Longitude and latitude (decimal degrees)                                                                 |
| VELU, VELV           | East and north components of velocity (cm/s)                                                             |
| VFLG                 | Vector validity (see below)                                                                              |
| RNGE                 | Distance from the antenna (km)                                                                           |
| BEAR                 | Bearing of the vector (counter-clockwise from true north)                                                |
| VELO                 | Radial velocity (cm/s)                                                                                   |
| HEAD                 | Velocity direction (counter-clockwise from true north)                                                   |
| SPRC                 | Range cell                                                                                               |
| SPDC                 | Doppler bin                                                                                              |
| MSEL                 | Selected solution (1 = single, 2 = dual 1, 3 = dual 2)                                                     |
| MSA1                 | Bearing of the single solution (1440 if invalid)                                                         |
| MDA1, MDA2           | Bearing of the first and second dual solutions (1440 if invalid)                                         |
| MEGR                 | Ratio of the first and second eigenvalue                                                                 |
| MPKR                 | Signal power ratio                                                                                       |
| MOFR                 | Off-diagonal ratio                                                                                       |
| MP13                 | Phase between antennas 1 and 3                                                                           |
| MP23                 | Phase between antennas 2 and 3                                                                           |
| MSP1                 | Signal power of the single solution (dB)                                                                 |
| MDP1, MDP2           | Signal power of the first and second dual solutions (dB)                                                 |
| MSW1                 | 3-dB width of the DOA function below the peak for the single solution                                     |
| MDW1, MDW2           | 3-dB width of the DOA function below the peak for the first and second dual solutions                      |
| MSR1                 | Value of DOA function at the peak of the single solution                                                 |
| MDR1                 | Value of DOA function at the peak of the first dual solution                                              |
| MDR2                 | Value of DOA function at the peak of the second dual solution                                              |
| MA1S, MA2S, MA3S     | SNR of the self-spectrum of antennas 1, 2, and 3 for this Doppler bin                                      |
| MEI1, MEI2, MEI3     | First, second, and third eigenvalue of the covariance matrix                                               |
| MDRJ                 | Reason for dual-solution rejection (see below)                                                           |
| PPFG, PWFG           | Flags for QARTOD tests 103 and 104 (1 = pass, 4 = fail)                                                   |

### VFLG

The value is composed as a sum of powers of 2, representing a binary number where each bit indicates a specific flag.

| Value   | Description                      |
| ------- | -------------------------------- |
| 0       | Valid vector                     |
| +64     | Low SNR                          |
| +4096   | Either PPFG or PWFG is not equal to 1 |

### MDRJ

This value is also composed as a sum of powers of 2, where each bit represents a flag.

| Value   | Description                                                                         |
| ------- | ----------------------------------------------------------------------------------- |
| 0       | Dual solution not rejected                                                          |
| +1      | P1 test failure                                                                     |
| +2      | P2 test failure                                                                     |
| +4      | P3 test failure                                                                     |
| +8      | The angle difference between both solutions is less than 20 degrees                 |
| +16     | Only one peak was found in the dual solution                                          |

