---
title: "Cross-Spectra Files"
output: rmarkdown::html_vignette
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Cross-Spectra}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(SeaSondeR)
```

## Introduction

Cross-spectra (CS) files contain a snapshot of the state of the ocean in cross-spectral format, calculated from the measurements of the three antennas. These data represent the reflected energy (self spectrum) for each detectable distance and Doppler velocity and also the cross product of the spectra (cross spectra) of the antennas relative to each other. These files are used to calculate radial velocity vectors and ocean state. The following information about the CS files is an extract from @SeaSonde-CSFileFormatV6, please see that document for more details on reading these types of files.

The files in this format are the raw cross spectra (CSQ) and the short time cross spectra (CSS). SeaSondeAcquisition creates the CSQs in the Spectra Series folder. SpectraAverager then reads the CSQs and produces CSS/CSAs in the SpectraToProcess folder. In a standard SeaSonde a CSS covers 15 minutes, with a file every 10 minutes. The CSA files are files produced by the SpectraSlider app and are an average of the corresponding CSS. The CSAs are used to process the second-order energy for obtaining wave height estimates. Both CSS and CSA are passed to the AnalyzeSpectra app to obtain radials and swell respectively.

They are binary files with the data in Big-Endian (Most Significant Byte first) which implies that in Intel we have to invert the order of the bytes to read the data. IEEE decimal values are used with single (4 bytes) and double (8 bytes) precision. Integer values are two's complement.

The file structure is:

-   Header section of variable size, which has been increasing with each version.

-   Data section, which can have two types of contents

    ## Header Section

Depending on the version of the file, it contains the following. Each higher version includes the info of the previous ones.

### Version 1

TimeStamp: seconds since January 1, 1904 of the local time of the station computer. For CSQs it is the start of data acquisition. For CSS it is the time of the center of the averaged data.

### Version 2

Data type: CSQ or CSS/CSA

### Version 3

Site name

### Version 4

Coverage in minutes of data. CSQ is usually about 5 minutes, CSS averages 15 minutes, CSA averages 60 minutes. Information about the initial frequency of each scan, Sweep rate, Sweep bandwidth, Sweep Freq direction. Number of Dopple cells (usually 512). Number of range cells (nominally 32 in CSQ and 31 in CSS/CSA). Index of the first range cell. Nominally 0 in CSQ, CSS usually uses 1 because SpectraAverager cuts the first range cell as useless. May take other values if the spectrum has been reduced to a certain range of interest. Cell distance in kilometers. The distance of a cell is the index of cell -1 + the index of the first cell \* the distance between cells.

### Version 5

Output interval in minutes. Number of active antennas. Number of antennas used in the spectra. Indicator of which antennas are in use.

### Version 6 (this section is variable in size and contains optional metadata).

TIME timestamp but with higher resolution. ZONE time use TOOL Name and version of the application(s) that processed the data in the file. ANTG Receive Antenna Gain corrections. It is an indicator of the power balance between the receiving antennas, which is used in SpectraPlotterMap to display the spectra. When applying antenna pattern or amplitude corrections they do not use this because they already take into account any kind of difference between the antennas. FOLS Radial/Elliptical First Order Lines to delineate the first order region. WOLS Wave Processing First Order Lones to delineate the first order region. RCVI a description of the receiver, including, among others, the receiver gain in dB. The conversion of the autospectrum to dBm is done as 10\*lob10(abs(autospectrum)) + receiver gain (if in the RCVI block, otherwise use -34.2 dB).

## Data section

This is a multidimensional array of auto-spectrum and cross-spectrum data. In the CSS and CSA, it also includes information on data quality.

# SeaSondeRCS Class

## Description

The SeaSondeRCS class is an R representation of the SeaSonde CS (Cross-spectra) files. This class facilitates the reading, manipulation and analysis of the data contained in these files, providing a consistent and easy-to-use structure for users working with SeaSonde data in R.

## Purpose

The main purpose of the SeaSondeRCS class is to provide a unified interface for working with CS files in R. This includes the ability to read CS files, access their metadata and data, and perform specific analyses related to SeaSonde data. By encapsulating this data in a specific class, the process of working with these files is simplified and greater consistency and reliability of analysis is ensured.

## Example of creation

To create an instance of the SeaSondeRCS class, you can use the following code in R:

```{r}

library(SeaSondeR)
# Create a SeaSondeRCS object form a CS file
CS_file <- "path/to/CSQ_o_CSS_o_CSA/file"

SeaSondeRCS_object <- seasonder_createSeaSondeRCS(x = CS_file)

```

In this example a SeaSondeRCS object is created from a CS file path.

## Error Management in SeaSondeRCS Class

The **`SeaSondeRCS`** class in R is equipped with a comprehensive error management system that ensures robust handling of issues that may arise during the creation and manipulation of SeaSonde cross-spectra (CS) files. This system is built using the **`rlang`** package, which provides detailed and structured error messages, making it easier for users to diagnose and address problems.

### Condition Classes

Several error classes have been defined to categorize common issues:

-   **`seasonder_CS_file_not_found_error`**: Triggered when the specified file path does not exist.

-   **`seasonder_CS_missing_nRange_nDoppler_error`**: Raised when the **`nRangeCells`** or **`nDopplerCells`** are not present in the header data.

-   **`seasonder_CS_data_structure_validation_error`**: Indicates a problem with the data structure of the CS data, as defined in the **`seasonder_validateCSDataStructure`** function.

-   **`seasonder_read_cs_file_error`**: An error class that indicates a general problem when attempting to read the SeaSonde CS file.

### Other condition classes

-   **`seasonder_cs_file_skipped`**: Condition indicating that the processing of a CS file was skipped due to an error.

# References
